---
title: "tut"
format: html
editor_options: 
  chunk_output_type: console
---

Get started with building a model in this R Markdown document that accompanies [Build a model](https://www.tidymodels.org/start/models/) tidymodels start article.

```{r}
library(tidymodels)  # for the parsnip package, along with the rest of tidymodels

# Helper packages
library(readr)       # for importing data
library(broom.mixed) # for converting bayesian models to tidy tibbles
library(dotwhisker)  # for visualizing regression results
```

## [The Sea Urchins Data](https://www.tidymodels.org/start/models/#data)

```{r}
urchins <-
  # Data were assembled for a tutorial 
  # at https://www.flutterbys.com.au/stats/tut/tut7.5a.html
  read_csv("https://tidymodels.org/start/models/urchins.csv") %>% 
  # Change the names to be a little more verbose
  setNames(c("food_regime", "initial_volume", "width")) %>% 
  # Factors are very helpful for modeling, so we convert one column
  mutate(food_regime = factor(food_regime, levels = c("Initial", "Low", "High")))
```

Look at the data:

```{r}
glimpse(urchins)
```

As a first step in modeling, it’s always a good idea to plot the data:

```{r message = FALSE}
ggplot(urchins,
       aes(x = initial_volume, 
           y = width, 
           group = food_regime, # redundant
           color = food_regime)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) +
  scale_color_viridis_d(option = "plasma", end = .7) + 
  theme_minimal()
```

We can see that urchins that were larger in volume at the start of the experiment tended to have wider sutures at the end, but the slopes of the lines look different so this effect may depend on the feeding regime condition.

## [Build and fit a model](https://www.tidymodels.org/start/models/#build-model)

A standard two-way analysis of variance (ANOVA) model makes sense for this dataset because we have both a continuous predictor and a categorical predictor. Since the slopes appear to be different for at least two of the feeding regimes, let’s build a model that allows for two-way interactions.

Specifying an R formula with our variables in this way:

```         
width ~ initial_volume * food_regime
```

allows our regression model depending on initial volume to have separate slopes and intercepts for each food regime.

For this kind of model, ordinary least squares is a good initial approach. With `tidymodels`, we start by specifying the functional form of the model that we want using the `parsnip` package. Since there is a numeric outcome and the model should be linear with slopes and intercepts, the model type is “linear regression”. We can declare this with:

```{r}
linear_reg()
```

Now that the type of model has been specified, we can think about a method for *fitting* or training the model, the model **engine**.

The default for `linear_reg()` is "lm" for ordinary least squares, as you can see above. We could set a non-default option instead:

```{r}
linear_reg() |> 
  set_engine("keras")
```

The [documentation page](https://parsnip.tidymodels.org/reference/linear_reg.html) for `linear_reg()` lists all the possible engines. We’ll save our model object using the default engine as `lm_mod`.

```{r}
lm_mod <- linear_reg()
```

From here, the model can be estimated or trained using the `fit()` function:

```{r}
lm_fit <- lm_mod |> 
  fit(width ~ initial_volume * food_regime, data = urchins)

lm_fit
```

Perhaps our analysis requires a description of the model parameter estimates and their statistical properties. Instead of `summary`, lets use `tidy`.

```{r}
tidy(lm_fit)
```

This kind of output can be used to generate a dot-and-whisker plot of our regression results using the `dotwhisker` package:

```{r}
tidy(lm_fit) |> 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
```

## [Use a model to predict](https://www.tidymodels.org/start/models/#predict-model)

This fitted object `lm_fit` has the `lm` model output built-in, which you can access with `lm_fit$fit`, but there are some benefits to using the fitted `parsnip` model object when it comes to predicting.

Suppose you want to make a plot of the mean body size for urchins that started the experiment with an initial volume of 20ml. To create such a graph, we start with some new example data that we will make predictions for, to show in our graph:

```{r}
new_points <- expand.grid(initial_volume = 20, 
                          food_regime = c("Initial", "Low", "High"))
new_points
```

To get our predicted results, we can use the `predict()` function to find the mean values at 20ml.

To communicate variability, we also need to find the predicted confidence intervals. If we had used `lm()` to fit the model directly, a few minutes of reading the [documentation page](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/predict.lm.html) for `predict.lm()` would explain how to do this.

However, if we decide to use a different model to estimate urchin size (spoiler: we will!), it is likely that a completely different syntax would be required.

Instead, with `tidymodels`, the types of predicted values are standardized so that we can use the same syntax to get these values.

First, let’s generate the mean body width values:

```{r}
mean_pred <- predict(lm_fit, new_data = new_points)
mean_pred
```

Get confidence intervals and plot:

```{r}
conf_int_pred <- predict(lm_fit, 
                         new_data = new_points, 
                         type = "conf_int")
conf_int_pred
```

```{r}
# Now combine: 
plot_data <- 
  new_points |> 
  bind_cols(mean_pred) |> 
  bind_cols(conf_int_pred)

# and plot:
ggplot(plot_data, aes(x = food_regime)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, 
                    ymax = .pred_upper),
                width = .2) + 
  labs(y = "urchin size")
```

## [Model with a different engine](https://www.tidymodels.org/start/models/#new-engine)

Wthe results would be different if the model were estimated using a Bayesian approach. In such an analysis, a [*prior distribution*](https://towardsdatascience.com/introduction-to-bayesian-linear-regression-e66e60791ea7/) needs to be declared for each model parameter that represents the possible values of the parameters (before being exposed to the observed data)

After some discussion, the group agrees that the priors should be bell-shaped but, since no one has any idea what the range of values should be, to take a conservative approach and make the priors wide using a Cauchy distribution (which is the same as a t-distribution with a single degree of freedom).

The [documentation](https://mc-stan.org/rstanarm/articles/priors.html) on the `rstanarm` package shows us that the `stan_glm()` function can be used to estimate this model, and that the function arguments that need to be specified are called `prior` and `prior_intercept.` It turns out that `linear_reg()` has a `stan` engine. Since these prior distribution arguments are specific to the Stan software, they are passed as arguments to `parsnip::set_engine()`. After that, the same exact `fit()` call is used:

```{r}
# set the prior distribution
prior_dist <- rstanarm::student_t(df = 1)

set.seed(123)

# make the parsnip model
bayes_mod <-   
  linear_reg() |> 
  set_engine("stan",
             prior_intercept = prior_dist,
             prior = prior_dist,
             refresh = 0)

# train the model
bayes_fit <- 
  bayes_mod |> 
  fit(width ~ initial_volume * food_regime, data = urchins)

print(bayes_fit, digits = 5)
```

This kind of Bayesian analysis (like many models) involves randomly generated numbers in its fitting procedure

To update the parameter table, the `tidy()` method is once again used:

```{r}
tidy(bayes_fit, conf.int = TRUE)
```

A goal of the \` packages is that the **interfaces to common tasks are standardized** (as seen in the [`tidy()`](https://generics.r-lib.org/reference/tidy.html) results above). The same is true for getting predictions; we can use the same code even though the underlying packages use very different syntax:

```{r}
bayes_plot_data <- 
  new_points |> 
  bind_cols(predict(bayes_fit, new_data = new_points)) |> 
  bind_cols(predict(bayes_fit, new_data = new_points, type = "conf_int"))

ggplot(bayes_plot_data, aes(x = food_regime)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, ymax = .pred_upper), width = .2) + 
  labs(y = "predicted urchin size") + 
  ggtitle("Bayesian model with t(1) prior distribution")
```

This isn’t very different from the non-Bayesian results (except in interpretation).

## Why does it work that way?

The extra step of defining the model using a function like [`linear_reg()`](https://parsnip.tidymodels.org/reference/linear_reg.html) might seem superfluous since a call to [`lm()`](https://rdrr.io/r/stats/lm.html) is much more succinct. However, the problem with standard modeling functions is that they don’t separate what you want to do from the execution. For example, the process of executing a formula has to happen repeatedly across model calls even when the formula does not change; we can’t recycle those computations.

Also, using the `tidymodels` framework, we can do some interesting things by incrementally creating a model (instead of using single function call). [Model tuning](https://www.tidymodels.org/start/tuning/) with tidymodels uses the specification of the model to declare what parts of the model should be tuned. That would be very difficult to do if [`linear_reg()`](https://parsnip.tidymodels.org/reference/linear_reg.html) immediately fit the model.

If you are familiar with the tidyverse, you may have noticed that our modeling code uses the base R pipe (`|>`). With dplyr and other tidyverse packages, the pipe works well because all of the functions take the *data* as the first argument. For example:

```{r}
urchins |> 
  group_by(food_regime) |> 
  summarize(med_vol = median(initial_volume))
```

whereas the modeling code uses the pipe to pass around the *model object*:

```{r}
bayes_mod |> 
  fit(width ~ initial_volume * food_regime, data = urchins)
```

This may seem jarring if you have used `dplyr` a lot, but it is extremely similar to how `ggplot2` operates:

```{r}
ggplot(urchins,
       aes(initial_volume, width)) +      # returns a ggplot object 
  geom_jitter() +                         # same
  geom_smooth(method = lm, se = FALSE) +  # same
  labs(x = "Volume", y = "Width") +       # etc
  theme_minimal()
```
