---
title: "tut"
format: html
editor_options: 
  chunk_output_type: console
---

# [Build a model](https://www.tidymodels.org/start/models/)

Get started with building a model in this R Markdown document that accompanies [Build a model](https://www.tidymodels.org/start/models/) `tidymodels` start article.

```{r}
library(tidymodels)  # for the parsnip package, along with the rest of tidymodels

# Helper packages
library(readr)       # for importing data
library(broom.mixed) # for converting bayesian models to tidy tibbles
library(dotwhisker)  # for visualizing regression results
```

## [The Sea Urchins Data](https://www.tidymodels.org/start/models/#data)

```{r}
urchins <-
  # Data were assembled for a tutorial 
  # at https://www.flutterbys.com.au/stats/tut/tut7.5a.html
  read_csv("https://tidymodels.org/start/models/urchins.csv") %>% 
  # Change the names to be a little more verbose
  setNames(c("food_regime", "initial_volume", "width")) %>% 
  # Factors are very helpful for modeling, so we convert one column
  mutate(food_regime = factor(food_regime, levels = c("Initial", "Low", "High")))
```

Look at the data:

```{r}
glimpse(urchins)
```

As a first step in modeling, it’s always a good idea to plot the data:

```{r message = FALSE}
ggplot(urchins,
       aes(x = initial_volume, 
           y = width, 
           group = food_regime, # redundant
           color = food_regime)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) +
  scale_color_viridis_d(option = "plasma", end = .7) + 
  theme_minimal()
```

We can see that urchins that were larger in volume at the start of the experiment tended to have wider sutures at the end, but the slopes of the lines look different so this effect may depend on the feeding regime condition.

## [Build and fit a model](https://www.tidymodels.org/start/models/#build-model)

A standard two-way analysis of variance (ANOVA) model makes sense for this dataset because we have both a continuous predictor and a categorical predictor. Since the slopes appear to be different for at least two of the feeding regimes, let’s build a model that allows for two-way interactions.

Specifying an R formula with our variables in this way:

```         
width ~ initial_volume * food_regime
```

allows our regression model depending on initial volume to have separate slopes and intercepts for each food regime.

For this kind of model, ordinary least squares is a good initial approach. With `tidymodels`, we start by specifying the functional form of the model that we want using the `parsnip` package. Since there is a numeric outcome and the model should be linear with slopes and intercepts, the model type is “linear regression”. We can declare this with:

```{r}
linear_reg()
```

Now that the type of model has been specified, we can think about a method for *fitting* or training the model, the model **engine**.

The default for `linear_reg()` is "lm" for ordinary least squares, as you can see above. We could set a non-default option instead:

```{r}
linear_reg() |> 
  set_engine("keras")
```

The [documentation page](https://parsnip.tidymodels.org/reference/linear_reg.html) for `linear_reg()` lists all the possible engines. We’ll save our model object using the default engine as `lm_mod`.

```{r}
lm_mod <- linear_reg()
```

From here, the model can be estimated or trained using the `fit()` function:

```{r}
lm_fit <- lm_mod |> 
  fit(width ~ initial_volume * food_regime, data = urchins)

lm_fit
```

Perhaps our analysis requires a description of the model parameter estimates and their statistical properties. Instead of `summary`, lets use `tidy`.

```{r}
tidy(lm_fit)
```

This kind of output can be used to generate a dot-and-whisker plot of our regression results using the `dotwhisker` package:

```{r}
tidy(lm_fit) |> 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
```

## [Use a model to predict](https://www.tidymodels.org/start/models/#predict-model)

This fitted object `lm_fit` has the `lm` model output built-in, which you can access with `lm_fit$fit`, but there are some benefits to using the fitted `parsnip` model object when it comes to predicting.

Suppose you want to make a plot of the mean body size for urchins that started the experiment with an initial volume of 20ml. To create such a graph, we start with some new example data that we will make predictions for, to show in our graph:

```{r}
new_points <- expand.grid(initial_volume = 20, 
                          food_regime = c("Initial", "Low", "High"))
new_points
```

To get our predicted results, we can use the `predict()` function to find the mean values at 20ml.

To communicate variability, we also need to find the predicted confidence intervals. If we had used `lm()` to fit the model directly, a few minutes of reading the [documentation page](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/predict.lm.html) for `predict.lm()` would explain how to do this.

However, if we decide to use a different model to estimate urchin size (spoiler: we will!), it is likely that a completely different syntax would be required.

Instead, with `tidymodels`, the types of predicted values are standardized so that we can use the same syntax to get these values.

First, let’s generate the mean body width values:

```{r}
mean_pred <- predict(lm_fit, new_data = new_points)
mean_pred
```

Get confidence intervals and plot:

```{r}
conf_int_pred <- predict(lm_fit, 
                         new_data = new_points, 
                         type = "conf_int")
conf_int_pred
```

```{r}
# Now combine: 
plot_data <- 
  new_points |> 
  bind_cols(mean_pred) |> 
  bind_cols(conf_int_pred)

# and plot:
ggplot(plot_data, aes(x = food_regime)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, 
                    ymax = .pred_upper),
                width = .2) + 
  labs(y = "urchin size")
```

## [Model with a different engine](https://www.tidymodels.org/start/models/#new-engine)

Wthe results would be different if the model were estimated using a Bayesian approach. In such an analysis, a [*prior distribution*](https://towardsdatascience.com/introduction-to-bayesian-linear-regression-e66e60791ea7/) needs to be declared for each model parameter that represents the possible values of the parameters (before being exposed to the observed data)

After some discussion, the group agrees that the priors should be bell-shaped but, since no one has any idea what the range of values should be, to take a conservative approach and make the priors wide using a Cauchy distribution (which is the same as a t-distribution with a single degree of freedom).

The [documentation](https://mc-stan.org/rstanarm/articles/priors.html) on the `rstanarm` package shows us that the `stan_glm()` function can be used to estimate this model, and that the function arguments that need to be specified are called `prior` and `prior_intercept.` It turns out that `linear_reg()` has a `stan` engine. Since these prior distribution arguments are specific to the Stan software, they are passed as arguments to `parsnip::set_engine()`. After that, the same exact `fit()` call is used:

```{r}
# set the prior distribution
prior_dist <- rstanarm::student_t(df = 1)

set.seed(123)

# make the parsnip model
bayes_mod <-   
  linear_reg() |> 
  set_engine("stan",
             prior_intercept = prior_dist,
             prior = prior_dist,
             refresh = 0)

# train the model
bayes_fit <- 
  bayes_mod |> 
  fit(width ~ initial_volume * food_regime, data = urchins)

print(bayes_fit, digits = 5)
```

This kind of Bayesian analysis (like many models) involves randomly generated numbers in its fitting procedure

To update the parameter table, the `tidy()` method is once again used:

```{r}
tidy(bayes_fit, conf.int = TRUE)
```

A goal of the \` packages is that the **interfaces to common tasks are standardized** (as seen in the [`tidy()`](https://generics.r-lib.org/reference/tidy.html) results above). The same is true for getting predictions; we can use the same code even though the underlying packages use very different syntax:

```{r}
bayes_plot_data <- 
  new_points |> 
  bind_cols(predict(bayes_fit, new_data = new_points)) |> 
  bind_cols(predict(bayes_fit, new_data = new_points, type = "conf_int"))

ggplot(bayes_plot_data, aes(x = food_regime)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, ymax = .pred_upper), width = .2) + 
  labs(y = "predicted urchin size") + 
  ggtitle("Bayesian model with t(1) prior distribution")
```

This isn’t very different from the non-Bayesian results (except in interpretation).

## Why does it work that way?

The extra step of defining the model using a function like [`linear_reg()`](https://parsnip.tidymodels.org/reference/linear_reg.html) might seem superfluous since a call to [`lm()`](https://rdrr.io/r/stats/lm.html) is much more succinct. However, the problem with standard modeling functions is that they don’t separate what you want to do from the execution. For example, the process of executing a formula has to happen repeatedly across model calls even when the formula does not change; we can’t recycle those computations.

Also, using the `tidymodels` framework, we can do some interesting things by incrementally creating a model (instead of using single function call). [Model tuning](https://www.tidymodels.org/start/tuning/) with tidymodels uses the specification of the model to declare what parts of the model should be tuned. That would be very difficult to do if [`linear_reg()`](https://parsnip.tidymodels.org/reference/linear_reg.html) immediately fit the model.

If you are familiar with the tidyverse, you may have noticed that our modeling code uses the base R pipe (`|>`). With dplyr and other tidyverse packages, the pipe works well because all of the functions take the *data* as the first argument. For example:

```{r}
urchins |> 
  group_by(food_regime) |> 
  summarize(med_vol = median(initial_volume))
```

whereas the modeling code uses the pipe to pass around the *model object*:

```{r}
bayes_mod |> 
  fit(width ~ initial_volume * food_regime, data = urchins)
```

This may seem jarring if you have used `dplyr` a lot, but it is extremely similar to how `ggplot2` operates:

```{r}
ggplot(urchins,
       aes(initial_volume, width)) +      # returns a ggplot object 
  geom_jitter() +                         # same
  geom_smooth(method = lm, se = FALSE) +  # same
  labs(x = "Volume", y = "Width") +       # etc
  theme_minimal()
```

# [Preprocess your data with recipes](https://www.tidymodels.org/start/recipes/)

## Introduction

In our [*Build a Model*](https://www.tidymodels.org/start/models/) article, we learned how to specify and train models with different engines using the [parsnip package](https://parsnip.tidymodels.org/). In this article, we’ll explore another `tidymodels` package, [recipes](https://recipes.tidymodels.org/), which is designed to help you preprocess your data *before* training your model. Recipes are built as a series of preprocessing steps, such as:

- converting qualitative predictors to indicator variables (also known as dummy variables),

- transforming data to be on a different scale (e.g., taking the logarithm of a variable),

- transforming whole groups of predictors together,

- extracting key features from raw variables (e.g., getting the day of the week out of a date variable),

and so on.

```{r}
library(tidymodels)      # for the recipes package, along with the rest of tidymodels

# Helper packages
library(nycflights13)    # for flight data
library(skimr)           # for variable summaries
```

## The New York City flight data

Let’s use the [nycflights13 data](https://github.com/hadley/nycflights13) to predict whether a plane arrives more than 30 minutes late. Let’s start by loading the data and making a few changes to the variables:

```{r}
set.seed(123)

flight_data <- flights |> 
  
  mutate(
    # Convert the arrival delay to a factor
    arr_delay = ifelse(arr_delay >= 30, "late", "on_time"),
    arr_delay = factor(arr_delay),
    # We will use the date (not date-time) in the recipe below
    date = lubridate::as_date(time_hour)
  ) |> 
  
  # Include the weather data
  inner_join(weather, by = c("origin", "time_hour")) |> 
  
  # Only retain the specific columns we will use
  select(dep_time, flight, origin, dest, air_time, distance, 
         carrier, date, arr_delay, time_hour) |> 
  
  # Exclude missing data
  na.omit() |> 
  
  # For creating models, it is better to have qualitative columns
  # encoded as factors (instead of character strings)
  mutate_if(is.character, as.factor)
```

We can see that about 16% of the flights in this data set arrived more than 30 minutes late.

```{r}
flight_data |> 
  count(arr_delay) |> 
  mutate(prop = n/sum(n))
```

Notice that the variable we created called `arr_delay` is a factor variable; **it is important that our outcome variable for training a logistic regression model is a factor.**

```{r}
glimpse(flight_data)
```

Second, there are two variables that we don’t want to use as predictors in our model, but that we would like to retain as identification variables that can be used to troubleshoot poorly predicted data points. These are `flight`, a numeric value, and `time_hour`, a date-time value.

Third, there are 104 flight destinations contained in `dest` and 16 distinct `carrier`s.

```{r}
flight_data |> 
  skimr::skim(dest, carrier) 
```

Because we’ll be using a simple logistic regression model, the variables `dest` and `carrier` will be converted to [dummy variables](https://bookdown.org/max/FES/creating-dummy-variables-for-unordered-categories.html). However, some of these values do not occur very frequently and this could complicate our analysis. We’ll discuss specific steps later in this article that we can add to our recipe to address this issue before modeling

## Data splitting

let’s split this single dataset into two: a *training* set and a *testing* set. We’ll keep most of the rows in the original dataset (subset chosen randomly) in the *training* set. The training data will be used to *fit* the model, and the *testing* set will be used to measure model performance.

To do this, we can use the [rsample](https://rsample.tidymodels.org/) package to create an object that contains the information on *how* to split the data, and then two more `rsample` functions to create data frames for the training and testing sets:

```{r}
# Fix the random numbers by setting the seed 
# This enables the analysis to be reproducible when random numbers are used 
set.seed(222)
# Put 3/4 of the data into the training set 
data_split <- initial_split(flight_data, prop = 3/4)

# Create data frames for the two sets:
train_data <- training(data_split)
test_data  <- testing(data_split)
```

## Create recipe and roles

Note that we can use a recipe to create a few new predictors and conduct some preprocessing required by the model.

Let’s initiate a new recipe for a simple logistic regression model:

```{r}
flights_rec <- 
  recipe(arr_delay ~ ., data = train_data) 
```

Now we can add [roles](https://recipes.tidymodels.org/reference/roles.html) to this recipe.

We can use the [`update_role()`](https://recipes.tidymodels.org/reference/roles.html) function to let recipes know that `flight` and `time_hour` are variables with a custom role that we called `"ID"` (a role can have any character value).

This tells the recipe to keep these two variables but not use them as either outcomes or predictors.

```{r}
flights_rec <- 
  recipe(arr_delay ~ ., data = train_data) |> 
  update_role(flight, time_hour, new_role = "ID") 
```

This can be convenient when, after the model is fit, we want to investigate some poorly predicted value.

To get the current set of variables and roles, use the [`summary()`](https://rdrr.io/r/base/summary.html) function:

```{r}
summary(flights_rec)
```

## Create features

What happens if we transform `date` column to `numeric`?

```{r}
flight_data %>% 
  distinct(date) %>% 
  mutate(numeric_date = as.numeric(date)) 
```

From `date` we can derive more meaningful features such as:

- the day of the week,
- the month, and
- whether or not the date corresponds to a holiday.

Add **steps** to your recipe to generate these features:

```{r}
flights_rec <- 
  recipe(arr_delay ~ ., data = train_data) |> 
  update_role(flight, time_hour, new_role = "ID") |> 
  step_date(date, features = c("dow", "month")) |>               
  step_holiday(date, 
               holidays = timeDate::listHolidays("US"), 
               keep_original_cols = FALSE)
```

What do each of these steps do?

- With [`step_date()`](https://recipes.tidymodels.org/reference/step_date.html), we created two new factor columns with the appropriate day of the week and the month.

- With [`step_holiday()`](https://recipes.tidymodels.org/reference/step_holiday.html), we created a binary variable indicating whether the current date is a holiday or not. The argument value of `timeDate::listHolidays("US")` uses the [timeDate package](https://cran.r-project.org/web/packages/timeDate/index.html) to list the 18 standard US holidays.

- With `keep_original_cols = FALSE`, we remove the original `date` variable since we no longer want it in the model. Many recipe steps that create new variables have this argument.

Because we plan to train a logistic regression model, we know that predictors will ultimately need to be numeric, as opposed to nominal data like strings and factor variables. For factors like `dest` and `origin`, [standard practice](https://bookdown.org/max/FES/creating-dummy-variables-for-unordered-categories.html) is to convert them into *dummy* or *indicator* variables to make them numeric.

But, unlike the standard model formula methods in R, a recipe **does not** automatically create these dummy variables for you; you’ll need to tell your recipe to add this step

```{r}
flights_rec <- 
  recipe(arr_delay ~ ., data = train_data) |> 
  update_role(flight, time_hour, new_role = "ID") |> 
  step_date(date, features = c("dow", "month")) |>               
  step_holiday(date, 
               holidays = timeDate::listHolidays("US"), 
               keep_original_cols = FALSE) |> 
  step_dummy(all_nominal_predictors())
```

Here, we did something different than before: instead of applying a step to an individual variable, we used [selectors](https://recipes.tidymodels.org/reference/selections.html) to apply this recipe step to several variables at once, [`all_nominal_predictors()`](https://recipes.tidymodels.org/reference/has_role.html). The [selector functions](https://recipes.tidymodels.org/reference/selections.html) can be combined to select intersections of variables.

We need one final step to add to our recipe. Since `carrier` and `dest` have some infrequently occurring factor values, it is possible that dummy variables might be created for values that don’t exist in the training set. For example, there is one destination that is only in the test set:

```{r}
test_data |> 
  distinct(dest) |> 
  anti_join(train_data)
```

When the recipe is applied to the training set, a column is made for LEX because the factor levels come from `flight_data` (not the training set), but this column will contain all zeros. This is a “zero-variance predictor” that has no information within the column.

[`step_zv()`](https://recipes.tidymodels.org/reference/step_zv.html) will remove columns from the data when the training set data have a single value, so it is added to the recipe *after* [`step_dummy()`](https://recipes.tidymodels.org/reference/step_dummy.html):

```{r}
flights_rec <- 
  recipe(arr_delay ~ ., data = train_data) |> 
  update_role(flight, time_hour, new_role = "ID") |> 
  step_date(date, features = c("dow", "month")) |>               
  step_holiday(date, 
               holidays = timeDate::listHolidays("US"), 
               keep_original_cols = FALSE) |> 
  step_dummy(all_nominal_predictors()) |> 
  step_zv(all_predictors())
```

Now we’ve created a *specification* of what should be done with the data. How do we use the recipe we made?

## Fit a model with a recipe

Let’s use logistic regression to model the flight data. As we saw in [*Build a Model*](https://www.tidymodels.org/start/models/), we start by [building a model specification](https://www.tidymodels.org/start/models/#build-model) using the parsnip package:

```{r}
lr_mod <- 
  logistic_reg() |> 
  set_engine("glm")
```

We will want to use our recipe across several steps as we train and test our model. We will:

1.  **Process the recipe using the training set**: This involves any estimation or calculations based on the training set. For our recipe, the training set will be used to determine which predictors should be converted to dummy variables and which predictors will have zero-variance in the training set, and should be slated for removal.

2.  **Apply the recipe to the training set**: We create the final predictor set on the training set.

3.  **Apply the recipe to the test set**: We create the final predictor set on the test set. Nothing is recomputed and no information from the test set is used here; the dummy variable and zero-variance results from the training set are applied to the test set.

To simplify this process, we can use a *model workflow*, which pairs a model and recipe together. We’ll use the [workflows package](https://workflows.tidymodels.org/) from tidymodels to bundle our parsnip model (`lr_mod`) with our recipe (`flights_rec`).

```{r}
flights_wflow <- 
  workflow() |> 
  add_model(lr_mod) |> 
  add_recipe(flights_rec)

flights_wflow
```

Now, there is a single function that can be used to prepare the recipe and train the model from the resulting predictors:

```{r}
flights_fit <- 
  flights_wflow |> 
  fit(data = train_data)
```

You may want to extract the model or recipe objects from the workflow. To do this, you can use the helper functions [`extract_fit_parsnip()`](https://hardhat.tidymodels.org/reference/hardhat-extract.html) and [`extract_recipe()`](https://hardhat.tidymodels.org/reference/hardhat-extract.html). For example, here we pull the fitted model object then use the [`broom::tidy()`](https://generics.r-lib.org/reference/tidy.html) function to get a tidy tibble of model coefficients:

```{r}
flights_fit |> 
  extract_fit_parsnip() |> 
  tidy()
```

## Use a trained workflow to predict

Our goal was to predict whether a plane arrives more than 30 minutes late. We have just:

1.  Built the model (`lr_mod`),

2.  Created a preprocessing recipe (`flights_rec`),

3.  Bundled the model and recipe (`flights_wflow`), and

4.  Trained our workflow using a single call to [`fit()`](https://generics.r-lib.org/reference/fit.html).

The next step is to use the trained workflow (`flights_fit`) to predict with the unseen test data, which we will do with a single call to [`predict()`](https://rdrr.io/r/stats/predict.html). The [`predict()`](https://rdrr.io/r/stats/predict.html) method applies the recipe to the new data, then passes them to the fitted model.

```{r}
predict(flights_fit, test_data)
```

Let’s say we want the predicted class probabilities for each flight instead. To return those, we can specify `type = "prob"` when we use [`predict()`](https://rdrr.io/r/stats/predict.html) or use [`augment()`](https://generics.r-lib.org/reference/augment.html) with the model plus test data to save them together:

```{r}
flights_aug <- 
  augment(flights_fit, test_data)

# The data look like: 
flights_aug |>
  select(arr_delay, time_hour, flight, .pred_class, .pred_on_time)
```

Now how will we evaluate the performance of our workflow? We can see from these first few rows that our model predicted these 5 on time flights correctly because the values of `.pred_on_time` are *p* \> .50. But we also know that we have 81,455 rows total to predict.

We would like to calculate a metric that tells how well our model predicted late arrivals, compared to the true status of our outcome variable, `arr_delay`.

Let’s use the area under the [ROC curve](https://bookdown.org/max/FES/measuring-performance.html#class-metrics) as our metric, computed using [`roc_curve()`](https://yardstick.tidymodels.org/reference/roc_curve.html) and [`roc_auc()`](https://yardstick.tidymodels.org/reference/roc_auc.html) from the [yardstick package](https://yardstick.tidymodels.org/).

To generate a ROC curve, we need the predicted class probabilities for `late` and `on_time`, which we just calculated in the code chunk above. We can create the ROC curve with these values, using [`roc_curve()`](https://yardstick.tidymodels.org/reference/roc_curve.html) and then piping to the [`autoplot()`](https://ggplot2.tidyverse.org/reference/autoplot.html) method:

```{r}
flights_aug |> 
  roc_curve(truth = arr_delay, .pred_late) |> 
  autoplot()
```

Similarly, [`roc_auc()`](https://yardstick.tidymodels.org/reference/roc_auc.html) estimates the area under the curve:

```{r}
flights_aug |> 
  roc_auc(truth = arr_delay, .pred_late)
```

Not too bad!

# Evaluate your model with resampling

## Introduction

Once we have a model trained, we need a way to measure how well that model predicts new data.

```{r}
library(tidymodels) # for the rsample package, along with the rest of tidymodels

# Helper packages
library(modeldata)  # for the cells data
library(ranger)
```

## The cell image data

Let’s use data from [Hill, LaPan, Li, and Haney (2007)](https://bmcbioinformatics.biomedcentral.com/articles/10.1186/1471-2105-8-340), available in the [modeldata package](https://cran.r-project.org/web/packages/modeldata/index.html), to predict cell image segmentation quality with resampling.

```{r}
data(cells, package = "modeldata")
cells
```

The main outcome variable of interest for us here is called `class`, which you can see is a factor. But before we jump into predicting the `class` variable, we need to understand it better.

### [Predicting image segmentation quality](https://www.tidymodels.org/start/resampling/#predicting-image-segmentation-quality)

Go to the link abvove to read more about this data.

A cell-based experiment might involve millions of cells so it is unfeasible to visually assess them all. Instead, a subsample can be created and these cells can be manually labeled by experts as either poorly segmented (`PS`) or well-segmented (`WS`). If we can predict these labels accurately, the larger data set can be improved by filtering out the cells most likely to be poorly segmented.

### Back to the cells data

The `cells` data has `class` labels for 2019 cells — each cell is labeled as either poorly segmented (`PS`) or well-segmented (`WS`). Each also has a total of 56 predictors based on automated image analysis measurements. For example, `avg_inten_ch_1` is the mean intensity of the data contained in the nucleus, `area_ch_1` is the total size of the cell, and so on (some predictors are fairly arcane in nature).

```{r}
cells
```

The rates of the classes are somewhat imbalanced; there are more poorly segmented cells than well-segmented cells:

```{r}
cells |> 
  count(class) |> 
  mutate(prop = n/sum(n))
```

## Data splitting

- The *training set* is used to estimate parameters, compare models and feature engineering techniques, tune models, etc.

- The *test set* is held in reserve until the end of the project, at which point there should only be one or two models under serious consideration. It is used as an unbiased source for measuring final model performance.

Since random sampling uses random numbers, it is important to set the random number seed. This ensures that the random numbers can be reproduced at a later time (if needed).

```{r}
set.seed(123)
cell_split <- initial_split(cells |> select(-case), 
                            strata = class)
```

Here we used the [`strata` argument](https://rsample.tidymodels.org/reference/initial_split.html), which conducts a stratified split. This ensures that, despite the imbalance we noticed in our `class` variable, our training and test data sets will keep roughly the same proportions of poorly and well-segmented cells as in the original data.

```{r}
cell_train <- training(cell_split)
cell_test  <- testing(cell_split)

nrow(cell_train)

nrow(cell_train)/nrow(cells)
```

```{r}
# training set proportions by class
cell_train |> 
  count(class) |> 
  mutate(prop = n/sum(n))

# test set proportions by class
cell_test |> 
  count(class) |> 
  mutate(prop = n/sum(n))
```

## Modeling

One of the benefits of a random forest model is that it is very low maintenance; it requires very little preprocessing of the data and the default parameters tend to give reasonable results. For that reason, we won’t create a recipe for the `cells` data.

Let's create the model specification for a [random forest model](https://en.wikipedia.org/wiki/Random_forest), using the `ranger` engine and setting the mode to `classification`.

```{r}
rf_mod <- 
  rand_forest(trees = 1000) |> 
  set_engine("ranger") |> 
  set_mode("classification")
```

Starting with this parsnip model object, the [`fit()`](https://generics.r-lib.org/reference/fit.html) function can be used with a model formula. Since random forest models use random numbers, we again set the seed prior to computing:

```{r}
set.seed(234)
rf_fit <- 
  rf_mod |> 
  fit(class ~ ., data = cell_train)

rf_fit
```

This new `rf_fit` object is our fitted model, trained on our training data set.

## Estimating performance

During a modeling project, we might create a variety of different models. To choose between them, we need to consider how well these models do, as measured by some performance statistics. In our example in this article, some options we could use are:

- the area under the Receiver Operating Characteristic (ROC) curve, and

- overall classification accuracy.

The ROC curve uses the class probability estimates to give us a sense of performance across the entire set of potential probability cutoffs. Overall accuracy uses the hard class predictions to measure performance. The hard class predictions tell us whether our model predicted `PS` or `WS` for each cell. But, behind those predictions, the model is actually estimating a probability. A simple 50% probability cutoff is used to categorize a cell as poorly segmented.

The [yardstick package](https://yardstick.tidymodels.org/) has functions for computing both of these measures called [`roc_auc()`](https://yardstick.tidymodels.org/reference/roc_auc.html) and [`accuracy()`](https://yardstick.tidymodels.org/reference/accuracy.html).

At first glance, it might seem like a good idea to use the training set data to compute these statistics. (This is actually a very bad idea.) Let’s see what happens if we try this. To evaluate performance based on the training set, we call the [`predict()`](https://rdrr.io/r/stats/predict.html) method to get both types of predictions (i.e. probabilities and hard class predictions).

```{r}
rf_training_pred <- 
  predict(rf_fit, cell_train) |> 
  bind_cols(predict(rf_fit, cell_train, type = "prob")) |> 
  # Add the true outcome data back in
  bind_cols(cell_train |> 
              select(class))
```

Using the yardstick functions, this model has spectacular results, so spectacular that you might be starting to get suspicious:

```{r}
rf_training_pred |>                # training set predictions
  roc_auc(truth = class, .pred_PS)

rf_training_pred |>                # training set predictions
  accuracy(truth = class, .pred_class)
```

Now that we have this model with exceptional performance, we proceed to the test set. Unfortunately, we discover that, although our results aren’t bad, they are certainly worse than what we initially thought based on predicting the training set:

```{r}
rf_testing_pred <- 
  predict(rf_fit, cell_test) |> 
  bind_cols(predict(rf_fit, cell_test, type = "prob")) |> 
  bind_cols(cell_test |> select(class))
```

```{r}
rf_testing_pred |>                   # test set predictions
  roc_auc(truth = class, .pred_PS)

rf_testing_pred |>                   # test set predictions
  accuracy(truth = class, .pred_class)
```

### What happened here?

There are several reasons why training set statistics like the ones shown in this section can be unrealistically optimistic:

- Models like random forests, neural networks, and other black-box methods can essentially memorize the training set. Re-predicting that same set should always result in nearly perfect results.

- The training set does not have the capacity to be a good arbiter of performance. It is not an independent piece of information; predicting the training set can only reflect what the model already knows.

To understand that second point better, think about an analogy from teaching. Suppose you give a class a test, then give them the answers, then provide the same test. The student scores on the *second* test do not accurately reflect what they know about the subject; these scores would probably be higher than their results on the first test.

## [Resampling to the rescue](https://www.tidymodels.org/start/resampling/#resampling)

Resampling methods, such as cross-validation and the bootstrap, are empirical simulation systems. They create a series of data sets similar to the training/testing split discussed previously; a subset of the data are used for creating the model and a different subset is used to measure performance. Resampling is always used with the *training set.*

Read the link above to see how their experiment played out.

## Fit a model with resampling

To generate these results, the first step is to create a resampling object using rsample. There are [several resampling methods](https://rsample.tidymodels.org/reference/index.html#section-resampling-methods) implemented in rsample; cross-validation folds can be created using [`vfold_cv()`](https://rsample.tidymodels.org/reference/vfold_cv.html):

```{r}
set.seed(345)
folds <- vfold_cv(cell_train, v = 10)
folds
```

The list column for `splits` contains the information on which rows belong in the analysis and assessment sets. There are functions that can be used to extract the individual resampled data called [`analysis()`](https://rsample.tidymodels.org/reference/as.data.frame.rsplit.html) and [`assessment()`](https://rsample.tidymodels.org/reference/as.data.frame.rsplit.html).

However, the tune package contains high-level functions that can do the required computations to resample a model for the purpose of measuring performance. You have several options for building an object for resampling:

- Resample a model specification preprocessed with a formula or [recipe](https://www.tidymodels.org/start/recipes/), or

- Resample a [`workflow()`](https://workflows.tidymodels.org/) that bundles together a model specification and formula/recipe.

For this example, let’s use a [`workflow()`](https://workflows.tidymodels.org/reference/workflow.html) that bundles together the random forest model and a formula, since we are not using a recipe. Whichever of these options you use, the syntax to [`fit_resamples()`](https://tune.tidymodels.org/reference/fit_resamples.html) is very similar to [`fit()`](https://generics.r-lib.org/reference/fit.html):

```{r}
rf_wf <- 
  workflow() |>
  add_model(rf_mod) |>
  add_formula(class ~ .)

set.seed(456)
rf_fit_rs <- 
  rf_wf |> 
  fit_resamples(folds)

rf_fit_rs
```

The results are similar to the `folds` results with some extra columns. The column `.metrics` contains the performance statistics created from the 10 assessment sets. These can be manually unnested but the tune package contains a number of simple functions that can extract these data:

```{r}
collect_metrics(rf_fit_rs)
```

Think about these values we now have for accuracy and AUC. These performance metrics are now more realistic (i.e. lower) than our ill-advised first attempt at computing performance metrics in the section above.

If we wanted to try different model types for this data set, we could more confidently compare performance metrics computed using resampling to choose between models. Also, remember that at the end of our project, we return to our test set to estimate final model performance. We have looked at this once already before we started using resampling, but let’s remind ourselves of the results:

```{r}
rf_testing_pred |>                   # test set predictions
  roc_auc(truth = class, .pred_PS)

rf_testing_pred |>                   # test set predictions
  accuracy(truth = class, .pred_class)
```

The performance metrics from the test set are much closer to the performance metrics computed using resampling than our first (“bad idea”) attempt. Resampling allows us to simulate how well our model will perform on new data, and the test set acts as the final, unbiased check for our model’s performance.
